Delocalization of Eigenvectors in Random Matrices

### 2025-03-20 Draft.tex
comments in red

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{amsthm}
\usepackage{mdframed}
\usepackage{graphicx}
\graphicspath{{./Photos/}}
\usepackage[english, russian]{babel}
\usepackage{stmaryrd}
\usepackage{hyperref}
\usepackage{setspace}
\pdfimageresolution=1200
\usepackage{hyperref}
\usepackage{doi}
\usepackage{enumitem}
\usepackage{ragged2e} 
\usepackage[numbers]{natbib}
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{array}
\usepackage{tikz}
\usepackage{lipsum}
\usetikzlibrary{arrows.meta, positioning, decorations.pathmorphing}
\usepackage[a4paper, margin=0.75in]{geometry}
\usepackage{afterpage}

\usepackage{titlesec}
%\titleformat{\section}[display]
  %{\normalfont\centering}
  %{\huge\bfseries ---\thesection---}  % Number with lines on both sides
  %{0.2ex}                            % Reduced space between number and title
 % {\Large\bfseries}                 
%  [\vspace{1ex}\titlerule\vspace{1.5ex}]

@@

\newcommand{\E}{\mathcal{E}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\usepackage{titlesec}

@@

\noindent In our analysis, we will utilize the flow method (\hyperref[sec:flow]{Section 3}), where the spectral parameter $z=E+i\eta$ of the resolvent of $H$ (defined in \hyperref[sec:prereq]{Section 2}) is varied at constant-speed in the upper-half plane $\mathbb{H}$ ($E\in\mathbb{R}$ and $\eta>0$) as the entries of $H$ are realized as Brownian motion \cite{bandSDE}.

@@

\noindent This is equivalent to $\mathbf{u}_\alpha$ being eigenvectors for the resolvent with $\frac{1}{\lambda_\alpha-z}$ as the eigenvalues. This also means that we have by spectral decomposition 
$$G(z) =\sum_{i=1}^N \frac{\mathbf{u}_\alpha\mathbf{u}_\alpha^*}{\lambda_\alpha-z}.$$
Now, recall that their empirical measure for $H$ is defined as \cite{dynamic}: $$\varrho_N(dx):=\frac{1}{N}\sum_{j=1}^N \delta(x-\lambda_j)dx,$$
Let $m(z)=m_N(z) = \int_\R \frac{\varrho_N(x)}{x-z}dx$ be its Stieltjes transform for $z = E+i\eta$, where $E\in \R$ and $\eta>0$. By the relationship we found above, we have the following  identity, which follows from the definition of the Dirac delta function: $$m(z)= \int_\R \frac{\varrho_N(x)}{x-z}dx = \int_\mathbb{R} \frac{1}{N}\frac{\sum_{j=1}^N \delta(x-\lambda_j)dx}{x-z}dx = \frac{1}{N}\sum_{j=1}^N\int_\R\frac{\delta(x-\lambda_j)}{x-z}dx =\frac{1}{N} \sum_{j=1}^N \frac{1}{\lambda_j-z} = \frac{1}{N}\text{Tr}(G(z))$$
From the moments method \cite{dynamic} we know that the Stieltjes transform $m_N(z)$ of the empirical distribution converges in probability to the Stieltjes transform $m_{SC} = \frac{1}{2}(-z+\sqrt{z^2-4})$ of the semicircle law as $N\rightarrow\infty$. This means that in the limit, with probability one the self-consistent equation holds: \begin{equation}
m(z) = -\frac{1}{z+m(z)} \Leftrightarrow m(z) + \frac{1}{m(z)} + z=0
\end{equation}
\subsection*{Useful identities}
Let us establish several facts and lemmas that we will need throughout our later work. We can begin by observing that: $$m(z)(m(z)+z) = \frac{1}{4}(-z+\sqrt{z^2-4})(z+\sqrt{z^2-4})=\frac{1}{4}(-z^2+z^2-4)=-1$$
$$\Rightarrow |m(z)| = \frac{1}{|m(z)+z|^{-1}}$$
Now, take the imaginary part of $(2)$: $$\text{Im}(m(z)) + \text{Im}\left(\frac{1}{m(z)}\right)+\eta=0$$
We can calculate the second term by simply expanding: $$\frac{1}{m(z)} = \frac{1}{\text{Re}(m(z))+\text{Im}(m(z))i} = \frac{\text{Re}(m(z))-\text{Im}(m(z))i}{|m(z)|^2}\Leftrightarrow \text{Im}\left(\frac{1}{m(z)}\right) = -\frac{\text{Im}(m(z))}{|m(z)|^2}$$
$$\Rightarrow\text{Im}(m(z)) + \text{Im}\left(\frac{1}{m(z)}\right)+\eta=\text{Im}(m(z)) -\frac{\text{Im}(m(z))}{|m(z)|^2}+\eta=0$$
$$\Rightarrow \text{Im}(m(z))\left(1-\frac{1}{|m(z)|^2}\right)=-\eta$$
And since $\text{Im}(m(z))>0$ and $\eta>0$, we must have $1-\frac{1}{|m(z)|^2}<0$, i.e $|m(z)|<1$, which is equivalent to $\text{Im }m \asymp 1$, where $\asymp$ means being bounded above and below up to a fixed, positive factor. Furthermore, let us rewrite the latter equation as the following two identities: \begin{equation*}1-|m(z)|^2 = \frac{\eta|m(z)|^2}{\text{Im}(m(z))}
\tag{3.1}
\end{equation*}
\begin{equation*}\text{Im}(m(z))= \frac{\eta|m(z)|^2}{1-|m(z)|^2 }
\tag{3.2}
\end{equation*}
Now, let $E\in [-10, 10]$ and $\eta\in (0, 10]$. First observe that $m(z)$ can never be zero, since it implies: $$m(z) = \frac{-z+\sqrt{z^2-4}}{2}=0 \Leftrightarrow \sqrt{z^2-4}=z$$
$$\Rightarrow z^2-4=z^2\Leftrightarrow -4=0$$ 
Then, clearly $\exists c>0$, s.t. $|m(z)|\geq c$.

@@

\subsection{Stopping time}
While the three terms we achieved in the previous section are a good start, we have non-linear powers w.r.t to term $\E_t(z)$ itself, which is what we are trying to bound. This means that we have to be more careful in our analysis. For this purpose, we will have to use a stopping time argument. More specifically, let us define:

>>>>

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{amsthm}
\usepackage{mdframed}
\usepackage{graphicx}
\graphicspath{{./Photos/}}
\usepackage[english, russian]{babel}
\usepackage{stmaryrd}
\usepackage{hyperref}
\usepackage{setspace}
\pdfimageresolution=1200
\usepackage{hyperref}
\usepackage{doi}
\usepackage{enumitem}
\usepackage{ragged2e} 
\usepackage[numbers]{natbib}
\usepackage[dvipsnames]{xcolor}
\usepackage{multirow}
\usepackage{array}
\usepackage{tikz}
\usepackage{lipsum}
\usetikzlibrary{arrows.meta, positioning, decorations.pathmorphing}
\usepackage[a4paper, margin=0.75in]{geometry}
\usepackage{afterpage}
\usepackage{titlesec}

\usepackage{titlesec}

@@

\newcommand{\E}{\mathcal{E}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\sto}{\text{stop}}
\newcommand{\im}{\text{Im}}

@@

\noindent In our analysis, we will utilize the flow method (\hyperref[sec:flow]{Section 3}), where the spectral parameter $z=E+i\eta$ of the resolvent of $H$ (defined in \hyperref[sec:prereq]{Section 2}) is varied at constant-speed in the upper-half plane $\mathbb{H}$ ($E\in\mathbb{R}$ and $\eta>0$) as the entries of $H$ are realized as Brownian motion \cite{bandSDE}.  \textcolor{red}{After writing up Quantum Diffusion, justify assumption }$\eta\gg W^{-3/4}$

@@

\noindent This is equivalent to $\mathbf{u}_\alpha$ being eigenvectors for the resolvent with $\frac{1}{\lambda_\alpha-z}$ as the eigenvalues. This also means that we have by spectral decomposition 
$$G(z) =\sum_{i=1}^N \frac{\mathbf{u}_\alpha\mathbf{u}_\alpha^*}{\lambda_\alpha-z}.$$
Now, recall that their empirical measure for $H$ is defined as \cite{dynamic}: $$\varrho_N(dx):=\frac{1}{N}\sum_{j=1}^N \delta(x-\lambda_j)dx,$$
Let $m(z)=m_N(z) = \int_\R \frac{\varrho_N(x)}{x-z}dx$ be its Stieltjes transform for $z = E+i\eta$, where $E\in \R$ and $\eta>0$. By the relationship we found above, we have the following  identity, which follows from the definition of the Dirac delta function: $$m(z)= \int_\R \frac{\varrho_N(x)}{x-z}dx = \int_\mathbb{R} \frac{1}{N}\frac{\sum_{j=1}^N \delta(x-\lambda_j)dx}{x-z}dx = \frac{1}{N}\sum_{j=1}^N\int_\R\frac{\delta(x-\lambda_j)}{x-z}dx =\frac{1}{N} \sum_{j=1}^N \frac{1}{\lambda_j-z} = \frac{1}{N}\text{Tr}(G(z))$$
From the moments method \cite{dynamic} we know that the Stieltjes transform $m_N(z)$ of the empirical distribution converges in probability to the Stieltjes transform $m_{SC} = \frac{1}{2}(-z+\sqrt{z^2-4})$ of the semicircle law as $N\rightarrow\infty$. This means that in the limit, with probability one the self-consistent equation holds: \begin{equation}
m(z) = -\frac{1}{z+m(z)} \Leftrightarrow m(z) + \frac{1}{m(z)} + z=0
\end{equation}
\subsection*{Useful identities}
Let us establish several facts and lemmas that we will need throughout our later work. We can begin by observing that: $$m(z)(m(z)+z) = \frac{1}{4}(-z+\sqrt{z^2-4})(z+\sqrt{z^2-4})=\frac{1}{4}(-z^2+z^2-4)=-1$$
$$\Rightarrow |m(z)| = \frac{1}{|m(z)+z|^{-1}}$$
Now, take the imaginary part of $(2)$: $$\text{Im}(m(z)) + \text{Im}\left(\frac{1}{m(z)}\right)+\eta=0$$
We can calculate the second term by simply expanding: $$\frac{1}{m(z)} = \frac{1}{\text{Re}(m(z))+\text{Im}(m(z))i} = \frac{\text{Re}(m(z))-\text{Im}(m(z))i}{|m(z)|^2}\Leftrightarrow \text{Im}\left(\frac{1}{m(z)}\right) = -\frac{\text{Im}(m(z))}{|m(z)|^2}$$
$$\Rightarrow\text{Im}(m(z)) + \text{Im}\left(\frac{1}{m(z)}\right)+\eta=\text{Im}(m(z)) -\frac{\text{Im}(m(z))}{|m(z)|^2}+\eta=0$$
$$\Rightarrow \text{Im}(m(z))\left(1-\frac{1}{|m(z)|^2}\right)=-\eta$$
And since $\text{Im}(m(z))>0$ and $\eta>0$, we must have $1-\frac{1}{|m(z)|^2}<0$, i.e $|m(z)|<1$, which is equivalent to $\text{Im }m \asymp 1$, where $\asymp$ means being bounded above and below up to a fixed, positive factor. Furthermore, let us rewrite the latter equation as the following two identities: \begin{equation*}1-|m(z)|^2 = \frac{\eta|m(z)|^2}{\text{Im}(m(z))}
\tag{3.1}
\end{equation*}
\begin{equation*}\text{Im}(m(z))= \frac{\eta|m(z)|^2}{1-|m(z)|^2 }
\tag{3.2}
\end{equation*}
Now, let $E\in [-10, 10]$ and $\eta\in (0, 10]$. First observe that $m(z)$ can never be zero, since it implies: $$m(z) = \frac{-z+\sqrt{z^2-4}}{2}=0 \Leftrightarrow \sqrt{z^2-4}=z$$
$$\Rightarrow z^2-4=z^2\Leftrightarrow -4=0$$ 
Then, clearly $\exists c>0$, s.t. $|m(z)|\geq c$. \textcolor{red}{Write lemma A.1 here or transfer the latter to the appendix}.

@@

\subsection{Stopping time}
While the three terms we achieved in the previous section are a good start, we have non-linear powers w.r.t to term $\E_t(z)$ itself, which is what we are trying to bound. This means that we have to be more careful in our analysis. For this purpose, we will have to use a stopping time argument \cite{bandSDE}. More specifically, for a fixed $\delta_\sto>0$ and $D\lesssim 1$ ($a\lesssim b\equiv a = O(b))$, independent of $N$, let us define the following:
\begin{equation*}
\tau_{\text{stop}, 1}=\inf \left\{s\geq 0: \max_{a, b} |\E_s(z)_{ab}|\geq W^{\delta_{\sto}}W^{-\frac{3}{4}}\cdot W^{-1}|\im w_s|^{-1}|\im w_s|^{-\frac{1}{2}}\right\}\wedge 1
\tag{3.2.1}
\end{equation*}

\begin{equation*}
\tau_{\text{stop}, 2}=\inf \left\{s\geq 0: \max_{a, b} 
\frac{|G_s(z)_{ab}-\delta_{ab}m(z)|^2}{(S^{1/2}T_s(z)S^{1/2})_{ab}+S_{ab}^{1/2}+W^{-D}}\geq W^{\delta_{\sto/10}}
\right\}\wedge 1
\tag{3.2.2}
\end{equation*}
\begin{equation*}
\tau_\sto = \tau_{\sto, 1}\wedge \tau_{\sto, 2}\tag{3.2.2}
\end{equation*}
